---
title: "Lab 06 - Truth-O-Meter"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(skimr)
```

```{r load-data, message = FALSE, eval = FALSE}

# Remove eval = FALSE or set it to TRUE once data is ready to be loaded

fact_checks_lie <- read_csv("data/fact_checks_lie.csv") # <<< you just created this .csv in the R scripts
fact_checks_truth <- read_csv("data/fact_checks_truth.csv") # <<< you just created this .csv in the R scripts

```

### Exercise 11

```{r add-new-variables-and-append, eval = FALSE}

fact_checks_lie <- fact_checks_lie %>%
  replace_with_function("___" = "___") # <<< create a new column that indicates these are lies

fact_checks_truth <- fact_checks_truth %>%
  replace_with_function("___" = "___") # <<< create a new column that indicates these are truths

fact_checks <- replace_with_function() # <<< combine the two data frames 

```

### Exercise 12

```{r separate-description-date-1, error = TRUE}

# Remove this text, and add your answer for Exercise 12 here.
```

### Exercise 13

Remove this text, and add your answer for Exercise 13 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 14

Remove this text, and add your answer for Exercise 13 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 15

Remove this text, and add your answer for Exercise 13 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 16

Remove this text, and add your answer for Exercise 13 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 17

Remove this text, and add your answer for Exercise 13 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.



