---
title: "Data classes"
subtitle: "<br><br> Behavioral Data Science in R"
author: "Derek Powell & Nicholas Duran"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(DT)
library(scales)
```

class: middle

## Data classes

We talked about *types* so far, next we'll introduce the concept of *classes*

-   Vectors are like Lego building blocks

| \- We stick them together to build more complicated constructs, e.g. *representations of data*                                                                                                                                                                  |
|-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| \- The **class** attribute relates to the S3 class of an object which determines its behavior 
| \- You don't need to worry about what S3 classes really mean, but you can read more about it [here](https://adv-r.hadley.nz/s3.html#s3-classes) if you're curious |

-   Examples: factors, dates, and data frames

---

## Factors

R uses factors to handle categorical variables, variables that have a fixed and known set of possible values

```{r}
x <- factor(c("BS", "MS", "PhD", "MS"))
x
```

---
                                               
## More on factors

We can think of factors like character (level labels) and an integer (level numbers) glued together

```{r}
glimpse(x)
as.integer(x)
```

---

## Dates

```{r}
y <- as.Date("2020-01-01")
y
typeof(y)
class(y)
```

---

## More on dates

We can think of dates like an integer (the number of days since the origin, 1 Jan 1970) and an integer (the origin) glued together

```{r}
as.integer(y)
as.integer(y) / 365 # roughly 50 yrs
```

---

## Data frames

We can think of data frames like vectors of equal length glued together

```{r}
df <- data.frame(x = 1:2, y = 3:4)
df
```


---

## Lists

Lists are a generic vector container vectors of any type can go in them

```{r}
l <- list(
  x = 1:4,
  y = c("hi", "hello", "jello"),
  z = c(TRUE, FALSE)
)
l
```

---

## Lists and data frames

-   A data frame is a special list containing vectors of equal length
-   When we use the `pull()` function, we extract a vector from the data frame
  - We can also use the `$` operator (e.g. `df$y`)

```{r}
df

df %>%
  pull(y)
```

---

class: middle

# Working with factors

---

```{r include=FALSE}
cat_lovers <- read_csv("data/cat-lovers.csv")
```

## Read data in as character strings

```{r}
glimpse(cat_lovers)
```

---

## But coerce when plotting

```{r out.width="60%"}
ggplot(cat_lovers, mapping = aes(x = handedness)) +
  geom_bar()
```

---

## Use forcats to manipulate factors

```{r out.width="55%"}
cat_lovers %>%
  mutate(handedness = fct_infreq(handedness)) %>% #<<
  ggplot(mapping = aes(x = handedness)) +
  geom_bar()
```

---

## Why factors?

- Factors are useful when you have true categorical data and you want to override the ordering of character vectors to improve display 
- They are also useful in modeling scenarios 
- The **forcats** package provides a suite of useful tools that solve common problems with factors

---

## Let's make a plot

```{r echo=FALSE, out.width="90%", fig.asp=0.4}
if (!file.exists("data/hotels")) {
  download.file("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv", "data/hotels.csv")
}
hotels <- readr::read_csv("data/hotels.csv")# %>% filter(arrival_date_year == 2016)

hotels %>%
  mutate(arrival_date_month = fct_relevel(arrival_date_month, month.name)) %>%
  group_by(hotel, arrival_date_month) %>%   # group by hotel type and arrival month
  summarise(mean_adr = mean(adr)) %>%       # calculate mean_adr for each group
  ggplot(aes(
    x = arrival_date_month,
    y = mean_adr,                           # y-axis is the mean_adr calculated above
    group = hotel,                          # group lines by hotel type
    color = hotel)                          # and color by hotel type
    ) +
  geom_line() +                             # use lines to represent data
  scale_y_continuous(labels = label_dollar()) +
  theme_minimal() +                         # use a minimal theme
  labs(x = "Arrival month",                 # customize labels
       y = "Mean ADR (average daily rate)",
       title = "Comparison of resort and city hotel prices across months",
       subtitle = "Resort hotel prices soar in the summer while city hotel prices remain\nrelatively constant throughout the year",
       color = "Hotel type") +
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
```

---

class: middle

# Working with dates

---

## Make a date

- **lubridate** is the tidyverse-friendly package that makes dealing with dates a little easier 
- It's not one of the *core* tidyverse packages, hence it's installed with `install.packages("tidyverse")` but it's not loaded with it, and needs to be explicitly loaded with `library(lubridate)`
- We're just going to scratch the surface of working with dates in R here...

---

## Hotels data

- Dataset of individual hotel bookings from 2015 to 2017

.small[
Calculate and visualize the number of bookings on any given arrival date.

```{r}
hotels %>%
  select(starts_with("arrival_"))
```
]

---

## Step 1. Construct dates

.small[
```{r}
hotels %>%
  mutate(
    arrival_date = 
      paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year)
      ) %>% 
  select(hotel, arrival_date)
```
]

---

## Step 2. Count bookings per date

```{r}
hotels %>%
  mutate(
    arrival_date = 
      paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year)
      ) %>% 
  select(hotel, arrival_date) %>% 
  count(arrival_date)
```

---

## Step 3. Visualise bookings per date

```{r, eval=F}
... %>% 
  ggplot(aes(x=arrival_date, y = n, group=1)) +
  geom_line() +
  theme(axis.text.x = element_text(angle=90, size=4))
```


```{r, echo=F}
hotels %>%
  mutate(
    arrival_date = 
      paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year)
      ) %>% 
  select(hotel, arrival_date) %>% 
  count(arrival_date) %>% 
  ggplot(aes(x=arrival_date, y = n, group=1)) +
  geom_line() +
  theme(axis.text.x = element_text(angle=90, size=4))
```


---

.hand[zooming in a bit...]

.question[ Why does the plot start with April 2016 when we know our data start in 2015?
And why does April 10th come after April 1st?]

```{r, echo=F}
hotels %>%
  mutate(
    arrival_date = 
      paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year)
      ) %>% 
  select(hotel, arrival_date) %>% 
  count(arrival_date) %>% 
  head(60) %>% 
  ggplot(aes(x=arrival_date, y = n, group=1)) +
  geom_line() +
  theme(axis.text.x = element_text(angle=90, hjust=1, size=6))
```




---

## Step 1. *REVISED* Construct dates "as dates"

```{r}
library(lubridate)

hotels %>%
  mutate(
    arrival_date_str = paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year),
    arrival_date = mdy(arrival_date_str) #<<
  ) %>% 
  select(hotel, arrival_date_str, arrival_date) 
```

---

## Step 2. Count bookings per date

```{r}
hotels %>%
  mutate(
    arrival_date_str = paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year),
    arrival_date = mdy(arrival_date_str)
  ) %>% 
  select(hotel, arrival_date_str, arrival_date) %>% 
  count(arrival_date) #<<
```


---

## Step 3a. Visualise bookings per date

```{r, eval=F}
... %>% 
  ggplot(aes(x=arrival_date, y = n, group=1)) +
  geom_line() 
```


```{r, echo=F}
hotels %>%
  mutate(
    arrival_date_str = paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year),
    arrival_date_str = parse_date_time(arrival_date_str, "mdY"),
    arrival_date = as_date(arrival_date_str)
  ) %>% 
  select(hotel, arrival_date_str, arrival_date) %>% 
  count(arrival_date) %>% 
  ggplot(aes(x=arrival_date, y = n, group=1)) +
  geom_line() 
```


---

## Step 3b. Visualize using a smooth curve

```{r, eval=F}
... %>% 
  ggplot(aes(x=arrival_date, y = n)) +
  geom_smooth() 
```


```{r, echo=F}
hotels %>%
  mutate(
    arrival_date_str = paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year),
    arrival_date_str = parse_date_time(arrival_date_str, "mdY"),
    arrival_date = as_date(arrival_date_str)
  ) %>% 
  select(hotel, arrival_date_str, arrival_date) %>% 
  count(arrival_date) %>% 
  ggplot(aes(x=arrival_date, y = n)) +
  # geom_point() +
  geom_smooth() 
  # theme(axis.text.x = element_text(angle=0, hjust=1, size=6))
```

