---
title: "Doing data science"
subtitle: "<br><br> Behavioral Data Science in R"
author: "School of Social and Behavioral Sciences"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    anchor_sections: FALSE
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

class: middle

# What's in a data analysis?

---

## Five core activities of data analysis

1. Stating and refining the question
1. Exploring the data
1. Building formal statistical models
1. Interpreting the results
1. Communicating the results

.footnote[
Roger D. Peng and Elizabeth Matsui. "The Art of Data Science." A Guide for Anyone Who Works with Data. Skybrude Consulting, LLC (2015).
]

---

class: middle

# Stating and refining the question

---

## Six types of questions

1. **Descriptive:** summarize a characteristic of a set of data
1. **Exploratory:** analyze to see if there are patterns, trends, or relationships between variables (hypothesis generating)
1. **Inferential:** analyze patterns, trends, or relationships in representative data from a population
1. **Predictive:** make predictions for individuals or groups of individuals
1. **Causal:** whether changing one factor will change another factor, on average, in a population
1. **Mechanistic:** explore "how" as opposed to whether

.footnote[
Jeffery T. Leek and Roger D. Peng. "What is the question?." Science 347.6228 (2015): 1314-1315.
]

---

## Ex: COVID-19 and Vitamin D

1. **Descriptive:** frequency of hospitalisations due to COVID-19 in a set of data collected from a group of individuals
--

1. **Exploratory:** examine relationships between a range of dietary factors and COVID-19 hospitalisations
--

1. **Inferential:** examine whether any relationship between taking Vitamin D supplements and COVID-19 hospitalisations found in the sample hold for the population at large

--
1. **Predictive:** what types of people will take Vitamin D supplements during the next year

--
1. **Causal:** whether people with COVID-19 who were randomly assigned to take Vitamin D supplements or those who were not are hospitalized

--
1. **Mechanistic:** how increased vitamin D intake leads to a reduction in the number of viral illnesses

---

## Questions to data science problems

- Do you have appropriate data to answer your question?
- Do you have information on confounding variables?
- Was the data you're working with collected in a way that introduces bias?

--

.question[
Suppose I want to estimate the average number of children in households in Edinburgh. I conduct a survey at an elementary school in Edinburgh and ask students at this elementary school how many children, including themselves, live in their house. Then, I take the average of the responses. Is this a biased or an unbiased estimate of the number of children in households in Edinburgh? If biased, will the value be an overestimate or underestimate?
]

---

class: middle

# Exploratory data analysis

---

## Checklist

- Formulate your question
- Read in your data
- Check the dimensions
- Look at the top and the bottom of your data
- Validate with at least one external data source
- Make a plot
- Try the easy solution first

---

## Formulate your question

- Consider scope:
  - Are air pollution levels higher on the east coast than on the west coast?
  - Are hourly ozone levels on average higher in New York City than they are in Los Angeles?
  - Do counties in the eastern United States have higher ozone levels than counties in the western United States?
- Most importantly: "Do I have the right data to answer this question?"

---


## Case study: Explanations in the wild!

> Why do some explanations strike people as highly satisfying while others, seemingly equally accurate, satisfy them less? We asked lay-people to generate and rate thousands of open-ended explanations in response to 'Why?' questions spanning multiple domains, and analyzed the properties of these explanations, to discover (1) what kinds of features are associated with greater explanation quality; (2) whether people can tell how good their explanations are; and (3) which cognitive traits predict the ability to generate good explanations. Our results support a pluralistic view of explanation, where satisfaction is best predicted by either functional or mechanistic content. Respondents were better able to judge how accurate their explanations were than how satisfying they were to others. Insight problem solving ability was the cognitive ability most strongly associated with the generation of satisfying explanations.

```{r, include=F}
eiw <- read_csv("../data/explanations-in-wild-s1-full.csv")
```

---

## Formulate your question

- Generate some possible questions
  - Do people give different kinds of explanations in different domains?
  - Are *good* explanations different in different domains?
  - What makes a satisfying explanation? How do mechanistic and functional explanations vary?
  - Is more more? Are longer explanations better or do people prefer shorter explanations?
  - Can people predict when their own explanations will be satisfying to others?  

--

- But as you do, ask yourself: "Do I have the right data to answer this question?"

---

## Figure out what is what

- The data for this paper <reference> are on the open science foundation website [osf.io](https://osf.io)
- As is often the case with open data (especially from preprints), we have to do some flipping back and forth between the paper and the data files online.

```{r, eval=FALSE, echo=TRUE}
eiw <- read_csv("explanations-in-wild.csv")
```


--

But once we do, we find there are 9 variables that look interesting: the question text, the explanation text, the domain of the question, and ratings for how accurate, satisfying, "causal", "mechanistic", "functional", and "general" each explanation is.

---

## Check the dimensions

```{r}
dim(eiw)
```

---

## Take a `glimpse()` at the data

```{r}
glimpse(eiw)
```


---

## First, just plot it

.hand[Do people give different explanations in different domains?]

.panelset[

.panel[.panel-name[Plot]
```{r ref.label="easy-solution", echo = FALSE}
```
]

.panel[.panel-name[Code]
```{r easy-solution, fig.show = "hide"}
eiw %>% 
  pivot_longer(
    c(functional, mechanism),
    names_to = "type",
    values_to = "rating"
  ) %>% 
    ggplot(aes(x=domain, y = rating, fill=type)) +
  geom_boxplot(width=.5)
```
]

]

---

## Then go deeper...

.small[
```{r facet-plot, echo=F}
eiw %>% 
  ggplot(aes(x = functional, y = satisfaction)) +
  geom_point() +
  geom_smooth(method="lm") +
  facet_wrap(~domain, labeller = label_wrap_gen(width=25)) +
  theme(aspect.ratio=1)
```
]
---

## Then go deeper... again

.small[
```{r hex-plot, echo=F}
eiw %>% 
  ggplot(aes(x = mechanism, y = satisfaction)) +
  geom_point() +
  geom_smooth(method="lm") +
  facet_wrap(~domain, labeller = label_wrap_gen(width=25)) +
  theme(aspect.ratio=1)
```
]

---

## And sometimes step back

Sometimes it helps to make visualizations of summaries as well.

- We notice that both mechanistic and functional explanations are more satisfying.
- New question: How much is everything correlated with everything else?

```{r heatmap, echo=F}
eiw %>% 
  mutate(
    log_text_length = log(stringr::str_length(text))
  ) %>% 
  select(satisfaction, accuracy, causation, mechanism, functional, log_text_length) %>% 
  cor() %>% 
  as_tibble(rownames = "var1") %>% 
  gather(var2, val, -var1) %>%  
  ggplot(aes(x=var1, y=var2, fill=val)) +
  geom_tile() +
  geom_text(aes(label=round(val,2))) +
  scale_x_discrete(limits=rev) +
  scale_fill_viridis_c() +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1), aspect.ratio=1) +
  labs(fill="corr.", x="", y="")
```

---

## Communicating for your audience

- Avoid: Jargon, uninterpreted results, lengthy output
- Pay attention to: Organization, presentation, flow
- Don't forget about: Code style, coding best practices, meaningful commits
- Be open to: Suggestions, feedback, taking (calculated) risks
