---
title: "Relationships among variables"
subtitle: "<br><br> Behavioral Data Science in R"
author: "School of Social and Behavioral Sciences"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(kableExtra)
library(dsbox)
```

class: middle

# Independence and dependence

---

## Conditional probability

**Notation**: $P(A | B)$: Probability of event A given event B

- What is the probability that it will be unseasonably warm tomorrow?
- What is the probability that it will be unseasonably warm tomorrow, given that it was unseasonably warm today?

---

## Independence

- If knowing event A happened tells you something about event B happening, or vice versa, then events A and B are **dependent**

  - $P(A|B) \neq P(A)$

- If not, they are said to be **independent**

  - $P(A | B) = P(A)$

---

class: middle

# Example: Climate survey

---

## Survey question

>A July 2019 YouGov survey asked 1633 GB and 1333 USA randomly selected adults
which of the following statements about the global environment best describes
their view:
>
>- The climate is changing and human activity is mainly responsible
>- The climate is changing and human activity is partly responsible, together with other factors
>- The climate is changing but human activity is not responsible at all
>- The climate is not changing

---

## Survey data

<br>

.small[
```{r echo=FALSE, message=FALSE}
yougov_climate_table <- read_csv("../data/yougov-climate.csv")

response_levels <- names(yougov_climate_table)[-1]

yougov_climate <- yougov_climate_table %>%
  pivot_longer(cols = -country, names_to = "response", values_to = "count") %>%
  uncount(count) %>%
  mutate(response = fct_relevel(response, response_levels))
```

```{r climate-table, echo=FALSE}
table(yougov_climate) %>%
  stats::addmargins() %>%
  knitr::kable(format = "html", booktabs = TRUE) %>%
  kableExtra::column_spec(column = 2:5, width = "0.5 in")
```
]

.footnote[
Source: [YouGov - International Climate Change Survey](https://d25d2506sfb94s.cloudfront.net/cumulus_uploads/document/epjj0nusce/YouGov%20-%20International%20climate%20change%20survey.pdf)
]

---

.question[
What percent of **all respondents** think the climate is changing and
human activity is mainly responsible?
]

.small[
```{r ref.label="climate-table", echo=FALSE}
```
]

--

```{r}
(all <- 1340 / 2966)
```


---

.question[
What percent of **GB respondents** think the climate is changing and
human activity is mainly responsible?
]

.small[
```{r ref.label="climate-table", echo=FALSE}
```
]

--

```{r}
(gb <- 833 / 1633)
```

---

.question[
What percent of **US respondents** think the climate is changing and
human activity is mainly responsible?
]

.small[
```{r ref.label="climate-table", echo=FALSE}
```
]

--

```{r}
(us <- 507 / 1333)
```

---

.question[
Based on the percentages we calculated, does there appear to be a relationship
between country and beliefs about climate change?
]

.pull-left[
```{r}
all
gb
us
```
]

---

class: middle

# Correlation

---

## Correlation is a kind of dependence


.pull-left[

So far we've primarily focused on _visualizing_ relationships between variables with plots like this:

In this plot, it looks like there is a positive association between X and Y: with higher values of X, we tend to see higher values of Y (and vice versa).

This **correlation** means that X and Y are **dependent**.
]

.pull-right[

```{r, echo=F, fig.height=10}
set.seed(12235)

tibble(
  X = rnorm(50),
  Y = .3*X + rnorm(50,0,.5)
) %>% 
  ggplot(aes(x=X, y=Y)) +
  geom_point() +
  theme(aspect.ratio=1)
```
]

---

## Correlation values
.small[
- Usually when someone refers to a correlation they refer to the "Pearson product-moment coefficient" or "Pearson correlation" (named for Karl Pearson).

- Typically denoted $\rho_{x,y}$, this is a number between -1 and 1 that determines the degree of linear association between two variables.
]

--

These plots show data with different $\rho_{x,y}$ values.

```{r, echo=F, warning=F}

make_2dcorr_samples <- function(n, rho){
  d <-  MASS::mvrnorm(n=n, mu=c(0, 0), Sigma=matrix(c(1, rho, rho, 1), nrow=2), empirical=TRUE) %>% 
    as_tibble() %>% 
    mutate(rho=rho)
  return(d)
}

cor_df <- bind_rows(map(c(.1, .25, .45, .75, .9), ~make_2dcorr_samples(100,.))) %>% 
  rename(X=V1, Y=V2)

cor_df %>% 
  rename(rho_val = rho) %>% 
  ggplot(aes(x=X, y=Y)) +
  geom_point() +
  facet_grid(~rho_val) +
  theme_bw() +
  theme(
    panel.grid=element_blank(),
    aspect.ratio=1,
    axis.text = element_blank()
    )
```

---

## Uncorrelated $\neq$ independent

.question[
Do these variables look independent?
]

```{r, echo=F}
set.seed(123435)
tibble(x=rnorm(100, 0,2),
       y = -x^2 + rnorm(100, 0,1)
       ) %>% 
  ggplot(aes(x=x, y=y)) +
  geom_point() +
  theme_bw(base_size=20) +
  theme(
    panel.grid=element_blank(),
    aspect.ratio=1,
    axis.text = element_blank()
    )
```

--
These variables are **uncorrelated**, but they are **not independent!**

---


