---
title: "Module 4 Code Along: Hotel bookings"
author:
  - Nicholas Duran, Derek Powell, Jessica Kosie^[Authored modifications to original code]
  - Mine Çetinkaya-Rundel^[Original author of exercise]
output:
  html_document:
    # theme: united
    theme: yeti
    highlight: pygments
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options: 
  chunk_output_type: console
---

```{r load-pkg, message = FALSE}

library(tidyverse)

```

### Understanding where our data comes from

We will continue working with the dataset used in last week's code-along. If you remember, this from the "R For Data Science" community podcast called Tidy Tuesday. The dataset is called "Hotel Bookings:" Learn more [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md)

```{r}

hotels <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv")

```

Let's visualize different aspects of hotel bookings on a few different timescales.

## Plot average daily rate (`adr`) by month

### Exercise 1. Plot average `adr` by arrival month for each hotel type. 

What are the steps do I need to take to solve this problem? Why not write some pseudocode?

```{r adr-by-month}

# We want to plot the average `adr` by hotel type and month. So, how do we do that?
# group the data by hotel type and month
# calculate the mean adr for each combination of hotel type and month
# what would be a good way to plot change over time?
# create a line graph where we can look at how the price changes across months - line graphs are good for plotting change over time
# use the line color to indicate hotel type

hotels %>% 
  group_by(hotel, arrival_date_month)  %>% 
  summarise(mean_adr = mean(adr)) %>%
  ggplot(aes(x=arrival_date_month, y=mean_adr, group=hotel, color=hotel)) +
    geom_line()

# what's wrong with this graph?

```

### Exercise 2. Fix ordering of the months (as factors)

We will want to use a function from the **forcats** package, see <https://forcats.tidyverse.org/reference/index.html> for inspiration and help.

```{r adr-by-month-ordered}

# copy and paste your code
# which of the functions in forcats do you think wold be useful here?

levels(hotels$arrival_date_month)
class(hotels$arrival_date_month)
# or look at tibble
hotels

# need to change it into a factor

hotels <- hotels %>%
  mutate(arrival_date_month_fct = fct_relevel(arrival_date_month, c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))) 

# now, let's look at the class and levels

class(hotels$arrival_date_month_fct)
levels(hotels$arrival_date_month_fct)

# R has some useful built-in constants, like `pi`, letters, LETTERS, month.name  
hotels_v2 <- hotels %>%
  mutate(arrival_date_month_fct = fct_relevel(arrival_date_month, month.name))

# now, let's look at the class and levels

class(hotels_v2$arrival_date_month_fct)
levels(hotels_v2$arrival_date_month_fct)

# now recreate your plot
hotels %>% 
  group_by(hotel, arrival_date_month) %>%
  summarise(mean_adr = mean(adr)) %>%
  ggplot(aes(x=arrival_date_month, y=mean_adr, group=hotel, color=hotel)) +
    geom_line()

# look at the x axis -- much better! but still not the most attractive plot.

```

### Exercise 3. Make it pretty

Totally up to you how to do this. Go wild. 

```{r adr-by-month-pretty}

# take ~10min with folks around you, try to make the plot pretty (labels, colors, theme)

hotels %>%
  group_by(hotel, arrival_date_month_fct) %>%
  summarise(mean_adr = mean(adr)) %>%
  ggplot(
    aes(x=arrival_date_month_fct, 
        y=mean_adr, 
        group=hotel, 
        color=hotel)
    ) +
    geom_line(linewidth = 1.5) +
  theme_classic() +
  labs(
    x = "Arrival month",
    y = "Mean average daily rate (ADR)",
    title = "City vs. Resort Hotels' ADR by Month",
    color = "Hotel Type"
  ) +
  scale_color_manual(values = c("purple", "forestgreen"))

```

**A final touch:** Change the y-axis label so the values are shown with dollar signs, e.g. $80 instead of 80.

We can use the `dollar_format()` function from the **scales** package, see <https://scales.r-lib.org/reference/index.html>.

```{r adr-by-month-dollars}

#copy and paste code and add dollar formatting

hotels %>%
  group_by(hotel, arrival_date_month_fct) %>%
  summarise(mean_adr = mean(adr)) %>%
  ggplot(
    aes(x=arrival_date_month_fct, 
        y=mean_adr, 
        group=hotel, 
        color=hotel)
    ) +
    geom_line(linewidth = 1.5) +
  theme_classic() +
  labs(
    x = "Arrival month",
    y = "Mean average daily rate (ADR)",
    title = "City vs. Resort Hotels' ADR by Month",
    color = "Hotel Type"
  ) +
  scale_color_manual(values = c("purple", "forestgreen")) +
  scale_y_continuous(labels = scales::dollar_format()) 

```

## Plot cancellations by day of week

### Exercise 4. Are travelers arriving different days (Sunday through Saturday) more likely to cancel their plans?

uh oh, no day of the week variable! we can use **lubridate** for this

```{r cancel-by-day-of-week}
library(lubridate)

hotels <- hotels %>%
  mutate(
    arrival_date_str = paste(arrival_date_month, arrival_date_day_of_month, arrival_date_year),
    arrival_date_mdy = parse_date_time(arrival_date_str, "mdy")
  ) 
  
hotels %>% 
  select(arrival_date_month, arrival_date_day_of_month, arrival_date_year, arrival_date_str, arrival_date_mdy)

hotels %>% 
  select(starts_with("arrival"))

days_of_week <- c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")

hotels <- hotels %>%
  mutate(
    arrival_date_weekday_number = wday(arrival_date_mdy),
    arrival_date_weekday = factor(arrival_date_weekday_number, labels=days_of_week)
  ) 

hotels %>% 
  select(arrival_date_mdy, arrival_date_weekday_number, arrival_date_weekday)

hotels %>%
  group_by(hotel, arrival_date_weekday) %>%
  summarize(mean_adr = mean(adr),
            n = n(),
            prop_cancelled = sum(is_canceled)/n()) %>%
  ggplot(aes(x=arrival_date_weekday, y=prop_cancelled, color=hotel, group=hotel)) +
  geom_line()

# what do you see from this graph? 

# add a title and sub-title that describe what you've found

hotels %>%
  group_by(hotel, arrival_date_weekday) %>%
  summarize(mean_adr = mean(adr),
            n = n(),
            prop_cancelled = sum(is_canceled)/n()) %>%
  ggplot(aes(x=arrival_date_weekday, y=prop_cancelled, color=hotel, group=hotel)) +
  geom_line() +
  labs(
    title = "Comparison of resort and city hotel cancellations across days of week",
    subtitle = "Cancellations peak on different days for different hotel types"
  )

```

