---
title: "Nobel winners - key file"
author: "Nicholas Duran; Jessica Kosie"
---

Code-Along File #1

Here we will practice reading and writing csv files.

First, let take a look at our file structure. What files and folders do we already have in our environment?

```{r load-packages, message=FALSE}

library(tidyverse)

```

Read in the `nobel.csv` file from the `data-raw/` folder.

Take a look at the data file. What kind of file is it? What function would you use to read in the file? What is the path to the file?

```{r load-data}

nobel <- read_csv("data-raw/nobel.csv")

```

Use the `glimpse()` function to take a look at the data. We have used this dataset before, so it should look familiar.

```{r}

glimpse(nobel)

```

Split the data into two (STEM and non-STEM):

  - Create a new data frame, `nobel_stem`, that filters for the STEM fields: (Physics, Medicine, Chemistry, and Economics).
Create another data frame, `nobel_nonstem`, that filters for the remaining fields.

**Hint:** Use the `%in%` operator when `filter()`ing.

```{r filtering-category}

# stem laureates
stem <- nobel %>%
  filter(category %in% c("Physics", "Chemistry", "Medicine", "Economics"))

# non-steam laureates
non_stem <- nobel %>%
  filter(!(category %in% c("Physics", "Chemistry", "Medicine", "Economics")))

```

Save the two data frames to `nobel-stem.csv` and `nobel_nonstem.csv`, respectively, in the  `data/` folder.

Hint: we use `read_csv()` to "read" in a data file. What might we use to "write" a data file? How do we tell R to put the file into the `data/` folder?

```{r save-output-csv}

write_csv(stem, "data/nobel_stem.csv")
write_csv(non_stem, "data/nobel_nonstem.csv")

```

Now, check out the `data/` folder to see if your files saved correctly.

