---
title: "HW 02 - Phoenix city employee salaries"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

## Packages

We'll use the **tidyverse** package for this analysis.
Run the following code in the Console to load this package.

```{r load-packages, message=FALSE, warning=FALSE, eval=TRUE}
library(tidyverse)
```

## Data

The data can be found in `data` folder and read in with the following command.

```{r}
phxsal <- read_csv("data/phx-salaries.csv")
```

## Exercises

### Exercise 1

Remove this text, and add your answer here, including the inline code.

### Exercise 2

Remove this text, and add your answer here.

### Exercise 3

Remove this text, and add your answer here.

### Exercise 4

```{r}
## Copy, paste, and complete the code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```

### Exercise 5

```{r}
library(ggridges)

# Remove this text, and add your answer here. Add code chunks as needed. 
```

### Exercise 6

Remove this text, and add your answer here. Add code chunks as needed. 
