---
title: "HW 07 - Smokers in Whickham"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

## Load packages and data

```{r eval = TRUE, message = FALSE, warning=F}
library(tidyverse) 
library(mosaicData) 
```

## Data

The dataset we'll use is called Whickham from the **mosaicData** package.
You can find out more about the dataset by inspecting their documentation, which you can access by running `?Whickham` in the Console or using the Help menu in RStudio to search for `Whickham`.

```{r, include=include_sol}
data("Whickham")
```

## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here including any code needed.

### Exercise 2

Remove this text, and add your answer for Exercise 2 here including any code needed.

### Exercise 3

Remove this text, and add your answer for Exercise 3 here including any code needed.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here including any code needed.


### Exercise 5

Remove this text, and add your answer for Exercise 5 here including any code needed.


### Exercise 6

Remove this text, and add your answer for Exercise 6 here including any code needed.


### Exercise 7

Remove this text, and add your answer for Exercise 7 here including any code needed.


### Exercise 8

Remove this text, and add your answer for Exercise 8 here including any code needed.
