---
title: "Homework Module 8"
author: "Homework Key"
output: html_notebook
---

## Warming up


```{r load-packages, message = FALSE, eval = TRUE}
library(tidyverse)
library(modelr)
```


```{r}
fb <- read_csv("data/FootballBrain.csv")
```


## Exercises

### Exercise 1

```{r}
fb <- fb %>% 
  filter(Group != "Control") %>% 
  drop_na(RT)
```

### Exercise 2

It's a character variable to start.

```{r}
fb <- fb %>% 
  mutate(Group = relevel(factor(Group), ref= "FBNoConcuss"))
```

### Exercise 3


```{r}
fb %>% 
  pivot_longer(
    c(LeftHipp, RT),
    names_to = "var"
    ) %>% 
  ggplot(aes(x=Group, y = value)) +
  geom_boxplot() +
  facet_wrap(~var, scales="free")

```

### Exercise 4

```{r}
fb %>% 
  group_by(Years) %>%
  mutate(concuss = if_else(Group == "FBNoConcuss", 0, 1)) %>%
  summarize(prop = mean(concuss), count = n()) %>% 
  ggplot(aes(x=Years, y = prop, size = count)) +
    geom_point()
```


### Exericse 5

Generally most responses here are fine. The boxplot might make it harder to think about as an "association". However, in the point plot, the points summarizing only 1 individual are a bit problematic.

### Exercise 6

add code for plots here

### Exercise 7

```{r}
summary(fit1 <- lm(Hipp ~ Years, fb))
```

### Exercise 8


```{r}
summary(fit2 <-  lm(Hipp ~ Years + Group, fb))
summary(fit3 <-  lm(Hipp ~ Years * Group, fb))
```

### Exericse 9

```{r}
AIC(fit1, fit2, fit3)
```

`fit2` is weakly preferred over `fit1`.

### Exercise 10

add these plots

### Exercise 11

```{r}
summary(fit4 <- lm(RT ~ Hipp, fb))
```

### Exercise 12

Pretty much any alternate hypothesis is fine.

### Exercise 13

```{r}
standardize <- function(x){
  (x - mean(x))/sd(x)
}

fb_standardized <- fb %>% 
  mutate_at(vars(Hipp, Years, RT), standardize)
```


### Exercise 14


```{r}

coef(fit1_s <- lm(RT ~ Hipp, fb_standardized))
coef(fit2_s <- lm(RT ~ Years, fb_standardized))  
coef(fit3_s <- lm(RT ~ Hipp + Years, fb_standardized))  
```

### Exercise 15

Explain in terms of standardized coefficients: a 1 SD increase in `Hipp` leads to a .65 SD increase in simple RT.

In `fit3_s` the `Years` coefficient becomes negligle but the `Hipp` coefficient is unaffected. `Hipp` predicts over and above `Years`, but `Years` does not predict over and above `Hipp`. A causal chain is more plausible.

