---
title: "Lab 05 - Sad plots"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(readxl) 
```

### Instructional staff employment trends

Load in data. Make sure you are changing "eval=FALSE" to "eval=TRUE" or just remove it altogether.

```{r label-me, eval=FALSE}
staff <- read_csv("labs/data/instructional-staff.csv")
```

### Exercise 1

Remove this text, and add your answer for Question 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

```{r label-me-1, eval=FALSE}

staff_long <- staff %>% 
  pivot_longer(2:ncol(staff), names_to = "year")

```

### Exercise 2

Remove this text, and add your answer for Question 2 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

```{r label-me-2}

staff_long %>% 
  ggplot(aes(x = year, y = value, color = faculty_type, group = faculty_type)) +
  geom_line() +
  theme_linedraw()

```

### Exercise 3

```{r label-me-3, fig.width = 10, fig.fullwidth = TRUE}

staff_long %>% 
  ggplot(aes(x = year, y = value, color = faculty_type, group = faculty_type)) +
  geom_line() +
  theme_linedraw() +
  labs(x = "Year", y = "Percentage of Hires", color = "Faculty Type")


```

### Exercise 4

```{r label-me-3, fig.width = 10, fig.fullwidth = TRUE}

```

### Exercise 5

```{r label-me-4, fig.width = 10, fig.fullwidth = TRUE}

staff_long %>% 
  mutate(hours = case_when(str_detect(faculty_type, "Part-Time") == TRUE ~ "Part-Time",
                           TRUE ~ "Other")) %>% 
  group_by(hours, year) %>% 
  summarise(avg_hires = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = avg_hires, color = hours, group = hours)) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = c("grey", "red")) +
  theme_linedraw() +
  labs(x = "Year", y = "Percentage of Hires", color = "Faculty Type")

staff_long %>% 
  mutate(hours = case_when(str_detect(faculty_type, "Part-Time") == TRUE ~ "Part-Time",
                           TRUE ~ "Other")) %>% 
  ggplot(aes(x = year, y = value, color = hours, group = faculty_type)) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = c("grey", "red")) +
  theme_linedraw() +
  labs(x = "Year", y = "Percentage of Hires", color = "Faculty Type")

staff_long %>% 
  mutate(hours = case_when(str_detect(faculty_type, "Part-Time") == TRUE ~ "Part-Time",
                           TRUE ~ "Other")) %>% 
  ggplot(aes(x = year, y = value, color = hours, group = faculty_type, linetype = faculty_type)) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = c("grey", "red")) +
  theme_linedraw() +
  labs(x = "Year", y = "Percentage of Hires", color = "Faculty Type")


```

Now, do Part 2: Fisheries. Add Exercise headings as needed.
```{r}

```

### Exercise 1

```{r label-me-5, fig.width = 10, fig.fullwidth = TRUE}

fisheries <- read_excel("labs/data/fisheries.xlsx")








```