---
title: "Lab 02 - Plastic waste"
# bibliography: references.bib
# biblio-style: apalike
# link-citations: yes
params: 
    SHOW_SOLS: TRUE # if i wanted to give solutions, could toggle back on. 
    TOGGLE: TRUE
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: cosmo
    number_sections: false
    include:
      in_header: assets/toggling.html
    #   after_body: https://uoepsy.github.io/assets/ccfooter.html
    css: assets/style-labs.css
---

```{r include = FALSE}
source('assets/setup.R')
# knitr::opts_chunk$set(eval = FALSE)
```

Plastic pollution is a major and growing problem, negatively affecting oceans and wildlife health.
[Our World in Data](https://ourworldindata.org/plastic-pollution) has a lot of great data at various levels including globally, per country, and over time.
For this lab we focus on data from 2010.

Additionally, National Geographic ran a data visualization communication contest on plastic waste as seen [here](https://www.nationalgeographic.org/funding-opportunities/innovation-challenges/plastic/dataviz/).


:::lo
**LEARNING OBJECTIVES**

-   Visualising numerical and categorical data and interpreting visualisations
-   Recreating visualizations
-   Getting more practice using with R, RStudio, Git, and GitHub
:::

# Getting started

Go to the course RStudio Cloud Workspace and locate the folder for the lab assignment, which should be named `lab-02-plastic-waste`.

First, open the R Markdown document `lab-02.Rmd` and Knit it.

Make sure it compiles without errors and you can preview the output within the Viewer Pane.  

The output should also be automatically saved as an `.html` file with the same name.

## Packages

We'll use the **tidyverse** package for this analysis.
Run the following code in the Console to load this package.

```{r load-packages, message=FALSE, warning=FALSE, eval=TRUE}
library(tidyverse)
```

## Data

The dataset for this assignment can be found as a csv file in the `data` folder of your repository.
You can read it in using the following.

```{r load-data, message=FALSE, eval=TRUE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

The variable descriptions are as follows:

-   `code`: 3 Letter country code
-   `entity`: Country name
-   `continent`: Continent name
-   `year`: Year
-   `gdp_per_cap`: GDP per capita constant 2011 international \$, rate
-   `plastic_waste_per_cap`: Amount of plastic waste per capita in kg/day
-   `mismanaged_plastic_waste_per_cap`: Amount of mismanaged plastic waste per capita in kg/day
-   `mismanaged_plastic_waste`: Tonnes of mismanaged plastic waste
-   `coastal_pop`: Number of individuals living on/near coast
-   `total_pop`: Total population according to Gapminder

---
# Warm up


`r qbegin(1)`
Recall that RStudio is divided into four panes. Without looking, can you name them all and briefly describe their purpose?


`r qend()`


`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`

The panes are called X, Y, Z, and Q!

`r solend()`

`r qbegin(2)`

-   Verify that the dataset has loaded into the Environment. How many observations are in the dataset? Clicking on the dataset in the Environment will allow you to inspect it more carefully. Alternatively, you can type `View(plastic_waste)` into the Console to inspect the data.


`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`

There are a total of `r nrow(plastic_waste)` observations in the dataset.

You can directly print the number of rows (observations) by typing `nrow(plastic_waste)` into the Console.

`r solend()`

`r qbegin(3)`
-   Have a quick look at the data and notice that there are cells taking the value `NA` -- what does this mean?

**Hint:** If you're not sure, run the command `?NA` which will lead you to the documentation.

`r qend()`


`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`

`NA` means "Not Available"--it is the typical code for a missing value.

`r solend()`

---

# Exercises



Let's start by taking a look at the distribution of plastic waste per capita in 2010.

```{r plastic_waste_per_cap-hist, eval=TRUE}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

One country stands out as an unusual observation at the top of the distribution.
One way of identifying this country is to filter the data for countries where plastic waste per capita is greater than 3.5 kg/person.

```{r plastic_waste_per_cap-max, eval=TRUE}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

Did you expect this result?

You might consider doing some research on Trinidad and Tobago to see why plastic waste per capita is so high there, or whether this is a data error.


`r qbegin(4)`
1.  Plot, using histograms, the distribution of plastic waste per capita faceted by continent. What can you say about how the continents compare to each other in terms of their plastic waste per capita?

**NOTE:** From this point onwards the plots and the output of the code are not displayed in the lab instructions, but you can and should the code and view the results yourself.
`r qend()`


Another way of visualizing numerical data is using density plots.

```{r plastic_waste_per_cap-dens}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_density()
```

And compare distributions across continents by colouring density curves by continent.

```{r plastic_waste_per_cap-dens-color}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent)) +
  geom_density()
```

The resulting plot may be a little difficult to read, so let's also fill the curves in with colours as well.

```{r plastic_waste_per_cap-dens-color-fill}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density()
```

The overlapping colours make it difficult to tell what's happening with the distributions in continents plotted first, and hence covered by continents plotted over them.
We can change the transparency level of the fill color to help with this.
The `alpha` argument takes values between 0 and 1: 0 is completely transparent and 1 is completely opaque.
There is no way to tell what value will work best, so you just need to try a few.

```{r plastic_waste_per_cap-dens-color-fill-alpha}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.7)
```

This still doesn't look great...


`r qbegin(5)`
1.  Recreate the density plots above using a different (lower) alpha level that works better for displaying the density curves for all continents.
`r qend()`


`r qbegin(6)`
2.  Describe why we defined the `color` and `fill` of the curves by mapping aesthetics of the plot but we defined the `alpha` level as a characteristic of the plotting geom.
`r qend()`


ðŸ§¶ âœ… â¬†ï¸  *Now is a good time to save and knit your document and confirm the outputs look as you expect them to.*

And yet another way to visualize this relationship is using side-by-side box plots.

```{r plastic_waste_per_cap-box}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

`r qbegin(7)`

Convert your side-by-side box plots from the previous task to [violin plots](http://ggplot2.tidyverse.org/reference/geom_violin.html). What do the violin plots reveal that box plots do not? What features are apparent in the box plots but not in the violin plots?

`r qend()`

`r qbegin(8)`
Visualize the relationship between plastic waste per capita and mismanaged plastic waste per capita using a scatterplot.

Describe the relationship.
    
**Remember:** We use `geom_point()` to make scatterplots.

`r qend()`

`r qbegin(9)`
Color the points in the scatterplot by continent.

Does there seem to be any clear distinctions between continents with respect to how plastic waste per capita and mismanaged plastic waste per capita are associated?
`r qend()`

`r qbegin(10)`
Visualize the relationship between plastic waste per capita and total population as well as plastic waste per capita and coastal population. You will need to make two separate plots.

Do either of these pairs of variables appear to be more strongly linearly associated?
`r qend()`

ðŸ§¶ âœ… â¬†ï¸  *Now is a good time to save and knit your document and confirm the outputs look as you expect them to.*

---

# Wrapping up

We don't expect you to complete all of the exercises within the hour reserved for the live workshop.
Ideally, you should have got to this point.
If you still have some time left, move on to the remaining exercises below.
If not, you should find a time to meet with your team and complete them after the workshop.
If you haven't had time to finish the exercises above, please ask for help before you leave!

`r qbegin(11)`

1.  Recreate the following plot, and interpret what you see in context of the data.

```{r echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE}
plastic_waste %>% 
  mutate(coastal_pop_prop = coastal_pop / total_pop) %>%
  filter(plastic_waste_per_cap < 3) %>%
  ggplot(aes(x = coastal_pop_prop, y = plastic_waste_per_cap, color = continent)) + 
    geom_point() +
    geom_smooth(color = "black") +
    scale_color_viridis_d() +
    labs(x = "Coastal population proportion (Coastal / total population)", 
         y = "Plastic waste per capita ", 
         color = "Continent",
         title = "Plastic waste vs. coastal population proportion",
         subtitle = "by continent") +
    theme_minimal()
```

**Hint:** The x-axis is a calculated variable. One country with plastic waste per capita over 3 kg/day has been filtered out. And the data are not only represented with points on the plot but also a smooth curve. The term "smooth" should help you [pick which geom to use](https://ggplot2.tidyverse.org/reference/index.html#section-geoms).

`r qend()`

# Submitting your work

Once you are happy with the final state of your work, change the YAML `output` from `html_output` to `pdf_output.` Now knit the document to produce a final PDF file. Upload the PDF to the Canvas assignment.  