---
title: "Lab 03 - Plastic Pollution"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---


## Packages

We'll use the **tidyverse** package for this analysis.
Run the following code in the Console to load this package.

```{r load-packages, message=FALSE, warning=FALSE, eval=TRUE}
library(tidyverse)
```

## Data

The dataset for this assignment can be found as a csv file in the `data` folder of your repository.

You can read it in using the following.

```{r load-data, message=FALSE, eval=TRUE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```


# Warmup

## Question 1 

Remove this text and enter your answer here.

## Exericse 2

Remove this text and enter your answer here.

## Exercise 3 

Remove this text and enter your answer here.

## Exercises

## Question 4 

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```

Remove this text and enter your answer here.

## Question 5

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```


## Question 6

Remove this text and enter your answer here.

## Question 7

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```

## Question 8

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```

## Question 9

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```

## Question 10

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```


# Bonus

## Question 11

```{r}
## Remove these comments and add your code for the exercise here
## Remember to label this code chunk (do not use spaces in the label name)
## Where appropriate, be sure to write your answers to questions as text below
```