---
title: "Lab 06 - Children's word learning"
# bibliography: references.bib
# biblio-style: apalike
# link-citations: yes
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/labs/")
  })
params: 
    SHOW_SOLS: FALSE # if i wanted to give solutions, could toggle back on. 
    TOGGLE: TRUE
---

```{r setup, include=FALSE}
source('assets/setup.R')

# :::red : use for pre-lab activity instructions
# :::lo : use for outlining the learning objectives
# :::green : use for tips on how to complete the task
# :::yellow : use for additional information on how to use R that does not directly have anything to do with the assignment tasks
# :::frame : use when want to set some additional instructions/code/examples apart from the assignment task when explaining something about how to use R

# :::blue : same color as question box background. Not sure how much this needs to be used. 
# :::statbox : looks a lot like yellow

```


< introduction to children's langauge learning >


:::lo
**LEARNING OBJECTIVES**

-   Visualize trends across time
-   Visualising relationships between variables
-   Discovering Simpson's paradox via visualizations
:::

# Getting started

Go to the course RStudio Cloud Workspace and locate the folder for the lab assignment, which should be named `lab-06-word-learning`.

First, open the R Markdown document `lab-06.Rmd` and Knit it.

Make sure it compiles without errors and you can preview the output within the Viewer Pane.  

The output should also be automatically saved as an `.html` file with the same name.

## Warm up

Before we introduce the data, let's warm up with a simple exercise.

-   Update the YAML, changing the author name to your name, and **knit** the document.

## Packages

Load any packages you need here. If you find you need another as you work through the lab, add it here and re-run the chunk.

```{r, include=F}
library(tidyverse)
```


## Data

The data for this lab is contained in a file called `wordbank-item-freq.csv`. Read that file in to a variable `wordbank`. When you've got it loaded, a `glimpse()` should look like this:

```{r, echo=F}
wordbank <- read_csv("../data/wordbank-item-freq.csv")

glimpse(wordbank)
```


The data for this lab originally comes from the [Wordbank](http://wordbank.stanford.edu/), an open database of children's vocabulary development run out of Stanford University. The data we'll be workign with is a summary of many parents' responses on the CDI for their children. The CDI was administered to many different parents/children when those children were at different ages. Based on these many responses, the average proportion of children who produce each word at each age was calculated.


This dataset has a few key variables:

- `item_definition`: The 
- `category`: The Wordbank category to which the word belongs
- `16:30`: These variables indicate the proportion of children who produce the word at each month from 16 mo (`16`) to 30 mo (`30`).
- `freq`: The number of times the word appears in Google's [Trillion Word Corpus](https://ai.googleblog.com/2006/08/all-our-n-gram-are-belong-to-you.html).

:::green
If you would like to explore more of the Wordbank data, you can do so with the `wordbankr` package. Also __note__: the words and word-frequencies were somewhat lazily joined by me---so don't use this data directly for any real research purposes without looking things over more carefully.
:::

# Exercises

Take turns answering the exercises.
<!-- If workign remotely through Zoom, make sure that the person taking the lead for an exercise is sharing their screen. -->

You don't have to switch at each exercise, you can find your a cadence that works for your team and stick to it.

## Get to know your data

---

Let's begin with some quick exploratory analyses to get you navigating around and familiar with this data.

`r qbegin(1)`
What are the 3 most commonly produced words by 18mo? Among the sounds, which is the one least likely to be produced at 30mo?
`r qend()`


`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`
```{r}

# Need to watch videos and read material to see if there is a suggested way of doing this in there. Add answers here. 

```
`r solend()`


---

`r qbegin(2)`

Let's plot the trajectory of children's word learning across time for some different words. For the "sounds" category, plot the trajectory of children's learning from 16mo to 30mo, with the proportion of children producing the word on the y-axis and age on the x-axis. Think about what kind of geom makes the most sense for this, and use color to differentiate the words. From your plot, which word do children tend to know best? Which word is learned the fastest from 16 to 25 months? Just for fun, speculate a bit as to why that word might be learned so quickly in that time period.
`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`

```{r, include=F, echo=FALSE}
wordbank %>% 
  filter(category=="sounds") %>%
  gather(month, prop, `16`:`30`) %>% 
  group_by(word) %>% 
  mutate(top_prop = max(prop), month=as.numeric(month)) %>% 
  ggplot(aes(x=month, y = prop, color=word, group=word)) +
  geom_line(size=1) +
  scale_color_viridis_d()


### 18-24mo is about when toddlers start running in addition to walking
### possibly leading to more (painful) falls
```
`r solend()`

`r qbegin(3)`

Now let's plot the trajectory of children's word learning across time for some different categories of words. As before, now for the categories in `c("sounds", "games_routines", "body_parts", "animals","clothing","people","pronouns","connecting_words","quantifiers","helping_verbs")`, proportion of children producing the word on the y-axis and age on the x-axis. Since there are a fair number of categories, you may want to facet or also use the `linetype` aesthetic to help you tell the lines apart.

`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`
```{r, include=F}
wordbank %>% 
  filter(category %in% c("sounds", "games_routines", "body_parts", "animals","clothing","people","pronouns","connecting_words","quantifiers","helping_verbs")) %>% 
  gather(month, prop, `16`:`30`) %>% 
  group_by(category, month) %>% 
  summarize(prop = mean(prop), top_prop = mean(ifelse(month==30, prop, NA), na.rm=T)) %>% 
  mutate(month=as.numeric(month)) %>% 
  ggplot(aes(x=month, y = prop, color=category, linetype=category, group=category)) +
  geom_line( size=1) +
  scale_color_viridis_d()
```
`r solend()`

<!-- ðŸ§¶ âœ… â¬†ï¸ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.* -->

---

## Word learning and word frequency

So far we have seen that some words (and word categories) seem easier for children to learn. One reason something might be easier to learn is if it is encountered more frequently. We can learn the streets we drive more frequently, the recipes we cook more often, and so forth. Could it be that children are faster to learn the words that appear in the english language more frequently?

The `freq` variable measures the frequency of the word in English, as the number of times the word appeared out of one trillion words collected by Google (1,024,908,267,229 to be exact).

`r qbegin(4)`
Real quick, what are the 10 most frequent words in this dataset? The 10 least frequent? 
`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`
```{r}

```
`r solend()`


`r qbegin(5)`
Make a visualization of the relationship between word frequency and the proportion of children who produce the word at two years of age. Dealing with the frequency of words directly can be a bit wonky, as the distribution of word frequency tends to be very skewed. So transform the frequency into log-frequency before plotting.  

To help you visualize, you may also want to add a linear fit to the plot with `geom_smooth(method="lm")`.
`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`
```{r, include=F}
wordbank %>% 
  filter(category!="time_words", category!="locations") %>%
  select(-downloaded, -item_id) %>%
  ggplot(aes(x=log(freq), y = `24`)) +
  geom_point() +
  geom_smooth(method="lm")
```
`r solend()`



`r qbegin(6)`
Is the relationship positive, negative, or basically non-existent? Is that at all surprising? What might explain it?
`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`
Negative, and yes surprising. Simpson's paradox
`r solend()`


`r qbegin(7)`
Make another visualization of the relationship between the log-frequency of each word and the proportion of children who produce the word at two years of age, but this time facet the plot by word category. Now what do you see?

Can you put into words the set of relationships observed between these two plots?

`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`
```{r, include=F}
wordbank %>% 
  select(-downloaded, -item_id) %>%
  ggplot(aes(x=log(freq), y = `24`)) +
  geom_point() +
  facet_wrap(~category) +
  geom_smooth(method="lm") 
```
`r solend()`

## Digging in qualitatively

Do you speak to a toddler the same way you speak to a friend? A professor?

< So it seems like some word categories are harder to learn in that even though they are common words to which children would be frquently exposed, they still do not learn to produce them themselves until later in development. This might be something about the word categories themselves (some researchers think so). But it also might be something about child-directed speech.

The word-frequency data came from Google Books data, based on the number of times words appeared in a collecton of about 8 million books. How might the words used in books differ from the words heard by toddlers?

:::green
For the curious, this [paper](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0213554) discusses some more detailed concerns about using these kinds of datasets. And this [github repository](https://github.com/hackerb9/gwordlist) has some telling discussion of using raw word frequency counts generated from Google N-gram corpora. It can be tempting to treat data as "truth"---but things are rarely so straightforward.
:::

```{r, echo=F}
library(ggrepel)

wordbank %>% 
  group_by(category) %>% 
  summarize(
    freq = mean(freq),
    prop = mean(`24`)
  ) %>% 
  ggplot(aes(x=log(freq), y = prop, label=category)) +
  geom_point() +
  geom_text_repel(segment.color="grey") +
  xlim(13, 23) +
  theme(panel.grid=element_blank()) +
  labs(x="Log-frequency", y="Avg. Proportion @ 24mo")
```


`r qbegin(8)`
Plotting the average proportion producing at 24mo and average log frequency of words for each category makes the unexpected negative relationship very clear. But as you look at this plot, are there any low frequency word categories children might hear relatively more often than they are written about in books?

`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`

`r solend()`


`r qbegin(9)`
What about the high-frequency word categories children struggle to learn like quantifiers and connecting words. What are the most frequent words in those categories? How often do you think children hear those words?

`r qend()`

`r solbegin(show=params$SHOW_SOLS, toggle=params$TOGGLE)`

```{r, include=F}
wordbank %>% 
  filter(category=="helping_verbs") %>% 
  arrange(desc(freq)) %>% 
  select(word, freq)
```


`r solend()`


---

Now go back through your write up to make sure you've answered all questions and all of your R chunks are properly labelled.

:::red
**SUBMITTING YOUR WORK**

Once you decide as a team that you're done with this lab, change the YAML `output` from `html_output` to `pdf_output.` Now knit the document to produce a final PDF file. Make sure all team members have access to the PDF and each person needs to upload the PDF as a Canvas assignment.   
:::
