---
title: "Homework Module 8"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

## Warming up


```{r load-packages, message = FALSE, eval = TRUE}
library(tidyverse)
library(modelr)
```


```{r load-data}
fb <- read_csv("data/FootballBrain.csv")
```


## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here including any code needed.


### Exercise 2

Remove this text, and add your answer for Exercise 2 here including any code needed.


### Exercise 3

Remove this text, and add your answer for Exercise 3 here including any code needed.


### Exercise 4

Remove this text, and add your answer for Exercise 4 here including any code needed.


### Exericse 5

Remove this text, and add your answer for Exercise 5 here including any code needed.


### Exercise 6

Remove this text, and add your answer for Exercise 6 here including any code needed.


### Exercise 7

Remove this text, and add your answer for Exercise 7 here including any code needed.


### Exercise 8

Remove this text, and add your answer for Exercise 8 here including any code needed.



### Exericse 9

Remove this text, and add your answer for Exercise 9 here including any code needed.


### Exercise 10

Remove this text, and add your answer for Exercise 10 here including any code needed.


### Exercise 11

Remove this text, and add your answer for Exercise 11 here including any code needed.


### Exercise 12

Remove this text, and add your answer for Exercise 12 here including any code needed.


### Exercise 13

Remove this text, and add your answer for Exercise 13 here including any code needed.



### Exercise 14

Remove this text, and add your answer for Exercise 14 here including any code needed.



### Exercise 15

Remove this text, and add your answer for Exercise 15 here including any code needed.


