---
title: "HW 05 - Legos"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
```

## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

```{r label-me, eval=FALSE}
# Just some dummy code. My solution required two pipes, you might do things different. Several ways to get to the correct solution (some easier than others).
dataset_you_imported %>% some_function(on_some_variable) %>% some_other_function(on_some_new_variable_generated_from_previous_step_in_pipe)
```

“In this sample, the first three most common names of purchasers are …”.

### Exercise 2

Remove this text, and add your answer for Exercise 2 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

### Exercise 7

...

### Exercise 8

...

### Exercise 9

...
