---
title: 'Module 5: Quiz Starter File'
output: html_document
date: "2026-09-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)

```

Quiz task:

Create the plot *exactly* as it is depicted on Canvas.

* You may use your handwritten notes and/or reference previous documents on Posit Cloud. Everything you need to create this graph is there, no internet searches are permitted. *

Using the stress_data dataset:

1. Create a new dataset that includes only weeks 2-6.
3. Relevel the condition factor so the legend appears in this order:
    No intervention, Expressive writing, Mindfulness, Cognitive reappraisal
4. Create a line graph showing mean stress score over time.
    - x-axis: week
    - y-axis: mean_stress_score
    - color: condition
5. Knit your file as a .pdf and upload it to Module 5: Quiz on Canvas.

Use the data below to create your plot.

```{r create-plot}

stress_data <- [READ IN DATA]

stress_data_cleaned <- [GET CORRECT ENTRIES AND CHANGE CONDITION LEVELS]

ggplot(stress_data_cleaned [ADD ARGUMENTS TO CREATE GRAPH])

```

