---
title: 'Module 5: Quiz'
output: html_document
date: "2026-09-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)

#data:
stress_data <- data.frame(
  week = rep(1:6, times = 4),
  condition = rep(
    c("Mindfulness", "No intervention", "Cognitive reappraisal", "Expressive writing"),
    each = 6
  ),
  mean_stress_score = c(
    7.2, 6.8, 6.2, 5.7, 5.1, 4.8,
    7.1, 7.0, 7.2, 7.1, 7.0, 7.1,
    7.3, 6.7, 6.0, 5.4, 4.9, 4.5,
    7.2, 6.9, 6.5, 6.1, 5.8, 5.5
  )
)
```

Quiz task:

* You may use your handwritten notes and/or reference previous documents on Posit Cloud. Everything you need to create this graph is there, no internet searches are permitted. *

Using the stress_data dataset:

1. Create a new dataset that includes only weeks 2-6.
3. Relevel the condition factor so the legend appears in this order:
    No intervention, Expressive writing, Mindfulness, Cognitive reappraisal
4. Create a line graph showing mean stress score over time.
    - x-axis: week
    - y-axis: mean_stress_score
    - color: condition
5. Knit your file as a .pdf and upload it to Module 5: Quiz on Canvas.

Use the data below to create your plot.

```{r create-plot}

stress_filtered <- stress_data %>%
  filter(week >= 2) %>%
  mutate(
    condition = fct_relevel(
      condition,
      "No intervention",
      "Expressive writing",
      "Mindfulness",
      "Cognitive reappraisal"
    )
  )

ggplot(
  stress_filtered,
  aes(x = week, y = mean_stress_score, color = condition)
) +
  geom_line(linewidth = 1) +
  labs(
    title = "Mean Stress Scores by Coping Strategy Condition",
    x = "Week",
    y = "Mean stress score",
    color = "Condition"
  ) 

```

