---
title: '&#32;'
output:
  html_document
---

```{r, include=F, dependson="setup.R"}
## Exam 1 specific

library(tidyverse)
library(nycflights13)

set.seed(42)

flights <- flights %>% 
  slice_sample(prop=.35)
```


The nycflights13 package contains information about all flights that departed from NYC (e.g. EWR, JFK and LGA) in 2013. The main data is in the `flights` data frame, but there are additional data sets which may help understand what causes delays, specifically:

- `weather`: hourly meteorological data for each airport
- `planes`: construction information about each plane
- `airports`: airport names and locations
- `airlines`: translation between two letter carrier codes and names

```{r}
nrow(flights)
```


```{r}
flights %>% 
  right_join(weather) %>% 
  nrow()
```

What can you tell about these datasets from this output?

A) The `flights` dataset contains rows that cannot be uniquely identified in the `weather` dataset
B) The `weather` dataset has some extra duplicate rows that should be filtered with `remove_duplicates()`.
C) There are some rows in the `weather` dataset representing locations and times where there were no flights.
D) The `flights` dataset contains extra rows that do not match with any rows in the `weather` dataset.
