---
title: '&#32;'
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r, include=F, dependson="setup.R"}

library(tidyverse)

```

```{r, eval=T, echo=F, message=F}

flights_temp <- tibble(
  year = runif(10000,2013,2013),
  month = round(runif(10000,1,12),0),
  day = round(runif(10000,1,31),0),
  # hour = round(runif(10000,1,24),0), 
  LGA = round(rnorm(10000, mean = 55, sd = 16),1), 
  JFK = round(rnorm(10000, mean = 56, sd = 17),1), 
  EWR = round(rnorm(10000, mean = 57, sd = 15),1),
  dep_delay = round(rnorm(10000, mean = 13, sd = 42),1)
) 

```

> Lets say we want to look at the relationship between flight delays and the temperature. Maybe colder days have more delays? But unfortunately, our temperature data for each airport (LGA, JFK, and EWR) is stored in separate columns, like below:

```{r, echo=T}
flights_temp
```

> In order to generate a plot with average temperature on the x-axis (like we did with wind gust a few questions back), we need to transform our data frame. What is the first thing we need to do?

```{r, echo=F, eval=F}
flights_temp %>% pivot_longer(cols = c("LGA", "JFK", "EWR"), names_to = c("origin"), values_to = "temp")
```

A) 
```
flights_temp %>% pivot_longer(cols = c("LGA", "JFK", "EWR"), names_to = c("origin"), values_to = "temp")
```

B)
```
flights_temp %>% mutate(origin = c("LGA", "JFK", "EWR"),
                        temp = values) 
```

C) 
```
flights_temp %>% mutate(temp = bind.rows(cols = c("LGA", "JFK", "EWR")) %>% arrange(month, day)
```

D)
```
flights_temp %>% bind.rows(cols = c("LGA", "JFK", "EWR")) %>% names_to = "origin" %>% arrange(month, day)
```




