---
title: '&#32;'
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r, include=F, dependson="setup.R"}

library(tidyverse)
library(nycflights13)

set.seed(42)

flights <- flights %>% 
  slice_sample(prop=.35)

```


> The last few questions have involved selecting a subset of the 19 variables from the `flights` data frame. As a reminder, below is what you get when you run `?flights` to pull up the documentation:

```{r, echo=F}
helpfile <- utils:::.getHelpFile(help(flights))
outfile <- tempfile(fileext = ".html")
tools:::Rd2HTML(helpfile, out =outfile)
rawHTML <- paste(readLines(outfile), collapse="\n")
knitr::asis_output(htmltools::htmlPreserve(rawHTML))
```

> What if we want to select only the following four variables (and just these four variables) from `flights`?
>
- dep_time
- dep_delay
- arr_time
- arr_delay

> From below, select all options that are valid ways for doing so 
(NOTE: THERE ARE MULTIPLE CORRECT RESPONSES)

A) `select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")`

B) `select(flights, starts_with("dep_"), starts_with("arr_"))`

C) `select(flights, ends_with("arr_time"), ends_with("dep_time"))`

D) `select(flights, contains("_time"), contains("arr_"))`





