---
title: '&#32;'
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r, include=F, dependson="setup.R"}

library(tidyverse)
library(nycflights13)

set.seed(42)

flights <- flights %>% 
  slice_sample(prop=.35)

```

> Lets continue working with the temperature data, where temperature for each airport (LGA, JFK, and EWR) is stored in separate columns: 

```{r, eval=T, echo=F, message=F}

flights_temp <- tibble(
  year = runif(10000,2013,2013),
  month = round(runif(10000,1,12),0),
  day = round(runif(10000,1,31),0),
  # hour = round(runif(10000,1,24),0), 
  LGA = round(rnorm(10000, mean = 55, sd = 16),1), 
  JFK = round(rnorm(10000, mean = 56, sd = 17),1), 
  EWR = round(rnorm(10000, mean = 57, sd = 15),1),
  dep_delay = round(rnorm(10000, mean = 13, sd = 42),1)
) 

```

```{r, echo=T}
flights_temp
```

> With the data in the format, one of the things we might want to do is convert temperature, which is currently in fahrenheit, to celsius.
>
> Here is the formula for converting fahrenheit (F) to celsius (C):

`C = (F-32)*5 / 9`

> In the following, please select the most __efficient__ way for converting each of the temperature columns from fahrenheit to celsius.

```{r, echo=FALSE, eval=FALSE}
fahren_to_cel <- function(F) {
  (F-32)*5 / 9
}
flights_temp %>% select(LGA:EWR) %>% map_dfr(., fahren_to_cel)
```

A) 
```
flights_temp$LGA = (flights_temp$LGA-32)*5 / 9
flights_temp$JFK = (flights_temp$JFK-32)*5 / 9
flights_temp$EWR = (flights_temp$EWR-32)*5 / 9
bind_cols(flights_temp$LGA, flights_temp$JFK, flights_temp$EWR)
```

B)
```
fahren_to_cel <- function(F) {
  flights_temp %>% select(LGA:EWR) %>% map_dfr(., (F-32)*5 / 9)
}
```

C) 
```
fahren_to_cel <- function(F) {
  (F-32)*5 / 9
}
flights_temp %>% select(LGA:EWR) %>% map_dfr(., fahren_to_cel)
```

D)
```
fahren_to_cel <- function(F) {
  (F-32)*5 / 9
}
bind_cols(fahren_to_cel(flights_temp$LGA), fahren_to_cel(flights_temp$JFK), fahren_to_cel(flights_temp$EWR))
```




