---
title: '&#32;'
output:
  html_document
---

```{r, include=F, dependson="setup.R"}
## Exam 1 specific

library(tidyverse)
library(nycflights13)

set.seed(42)

flights <- flights %>% 
  slice_sample(prop=.35)
```


The nycflights13 package contains information about all flights that departed from NYC (e.g. EWR, JFK and LGA) in 2013. The main data is in the `flights` data frame, but there are additional data sets which may help understand what causes delays, specifically:

- `weather`: hourly meteorological data for each airport
- `planes`: construction information about each plane
- `airports`: airport names and locations
- `airlines`: translation between two letter carrier codes and names

Here are some selected rows of the `flights` dataset:

```{r}
flights %>% 
  select(flight, month, day, distance, carrier) %>% 
  head()
```

And here is a preview of the `airlines` dataset:

```{r}
airlines
```

How could we create a new `flights` dataframe that includes the full airline company names rather than just the unfamiliar abbreviations?

A) `flights <- flights %>% filter(carrier %in% airlines$names)`
B) `flights <- left_join(flights, airlines, by = c("carrier"="carrier"))` 
C) `flights <- flights %>% right_join(airlines, by = c("carrier"="name"))` 
D) `flights <- flights %>% select(carrier, airlines::name)` 
