---
title: "LAB 01 - Parks and Recreation"
author: "Jessica Kosie"
date: "08/26/2025"
output:
  pdf_document: default
  html_document: default
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
```

This is a R Markdown document!

## Exercises

```{r load-data, message = FALSE}

parks_rec <- read_csv("data/parks_rec.csv")

```

### Exercise 1

There are \_\_\_ unique characters from the show in the dataset.

### Exercise 2

Remove this text, and add your answer for Exercise 2 here.

### Exercise 3

Remove this text, and add your answer for Exercise 3 here. (Hint: You might also want to click on the green arrow pointing right icon in the code chunk below to help with your answer)

```{r most-common-characters}
parks_rec %>%
  count(character, sort = TRUE)
```

### Exercise 4

```{r distinct-episodes-a-season}
# remove this comment and add the code for Exercise 4 here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.
