---
title: "Exploratory data analysis 1"
author: "Derek Powell"
output: github_document
---

```{r load-packages, include=FALSE}
library(tidyverse)
```


Questions:

* How much does happiness vary by country?
* Which countries are happiest? Which are least happy?
* Can you buy happiness? How does per capita GDP relate to happiness?
* What is the cost of happiness? How does happiness relate to ecological footprint?


### Load the Happy Planet Index data frame we will be using 

```{r}
happy <- read_csv("https://www.lock5stat.com/datasets3e/HappyPlanetIndex.csv") %>% 
  mutate(
    Region = case_when(
      Region == 1 ~ "Latin America",
      Region == 2 ~ "Western",
      Region == 3 ~ "Middle East",
      Region == 4 ~ "SubSaharan Africa",
      Region == 5 ~ "South Asia",
      Region == 6 ~ "East Asia",
      Region == 7 ~ "Former Soviet",
      TRUE ~ NA_character_
    ),
    HDI_group = ordered(case_when(
      HDI <= .55 ~ "Low",
      HDI < .699 ~ "Medium",
      HDI < .799 ~ "High",
      HDI >= .799 ~ "Very high"
    ), levels = c("Low","Medium","High","Very high")
  )
  )
```


### Glimpse at the data

```{r glimpse-starwars}
glimpse(happy)
```

### How much does happiness vary across countries?

```{r}
ggplot(happy, aes(x=Happiness)) +
  geom_histogram()
```

### Which are the happiest countries?

```{r}

```



### Modify the following plot to change the color of all points to `"pink"`.

```{r scatterplot}
ggplot(happy, 
       aes(x = GDPperCapita, y = Happiness, size=Population)) +
  geom_point()
```

### Add labels for title, x and y axes, and size of points. Uncomment to see the effect.

```{r scatterplot-labels}
ggplot(happy, 
       aes(x = GDPperCapita, y = Happiness, size=Population)) +
  geom_point(color = "#30509C") +
  labs(
    #title = "___",
    #x = "___",
    #y = "___",
    #___
    )
```

### How is the data organized? Make a bar plot of the distribution of `Region`

(A little bit of starter code is provided below, and the code chunk is set to not be evaluated with `eval = FALSE` because the current code in there is not valid code and hence the document wouldn't knit. Once you replace the code with valid code, set the chunk option to `eval = TRUE`, or remove the `eval` option altogether since it's set to `TRUE` by default.)

```{r barplot, eval = FALSE}
ggplot(happy, aes(x=HDI_group)) + geom_bar()
```

### Pick a single numerical variable and make a histogram of it.

(This time no starter code is provided, you're on your own!)

```{r histogram}

```

### Pick a numerical variable and a visualisation (you pick the type!) to visualise the relationship between this variable and `Region`. Along with your code and output, provide an interpretation of the visualisation.

```{r num-cat}
```

### Focus on the two categorical variables of `Region` and `HDI_group` and make a visualisation to visualise the relationship between them. Along with your code and output, provide an interpretation of the visualisation.

```{r cat-cat}
```

### Pick two numerical variables and to consider alongside the two categorical variables of `Region` and `HDI_group` and make a visualisation that incorportes all of them. Then, provide an interpretation with your answer.

```{r multi}
```